\class : Week \psetnum

### 2022-12-05 541_ism/week9/solutions.tex
ISM week 9 problems all done

\newcommand{\psetnum}{7}
\newcommand{\class}{ASTR 541 - Interstellar Medium}

@@

\question{\textbf{1. HI 21cm Radio Observations}}

\question{1a. On/off Line}
\answer{
    On-line means taking a measurement at the frequency the line (in this case, the HI 21cm line)\footnote{Not to be confused with ``online'', which is a term that the kids use when they're referring to this new-fangled interweb thing.}, whereas an off-line measurement is taken at a different frequency, outside of the line profile of the target line.

    \noindent Tobin put it nicely as on-line being like tuning into the 21.11cm FM radio station and off-line as listening to the static between stations.
    \begin{align}
        F_{\rm on-line} &= F_{\nu, 0}\, e^{-\tau_\nu},\qquad \text{a.k.a. ``You're listening to 21.11cm FM!''} \\
        F_{\rm off-line} &= F_{\nu, 0}\,,\qquad\,\,\,\,\,\quad \text{a.k.a. ``Hsssttttt''}
    \end{align}
}

\question{1b. Optical Depth Limit}
\answer{
    We are given that the change in flux density is below 1\%. We can use this to solve for what optical depth this gives.
    \begin{align}
        F_{\rm off-line} - F_{\rm on-line} &< 0.01 F_{\rm off-line}\\
        1 - e^{-\tau_\nu} &< 0.01\\
        e^{-\tau_\nu} &> 0.99\\
        \Aboxed{ \tau_\nu &> 0.01 }
    \end{align}
}

\question{1c. Temperature Limit}
\answer{
    The equation in the problem gives that
    \begin{equation}
        \tau_\nu = 0.552 \qty(\frac{100 \unit{K}}{T_{\rm spin}}) \qty(\frac{\dv{N_{\rm HI}}{\nu}}{\frac{10^{20} \unit{cm^{-2}}}{\unit{km}{s^{-1}}}}),
    \end{equation}
    which can be quite easily transformed to get $T_{\rm spin}$ as a function of optical depth
    \begin{equation}
        T_{\rm spin} = 55.2 \unit{K}\cdot \tau_\nu \qty(\frac{\dv{N_{\rm HI}}{\nu}}{\frac{10^{20} \unit{cm^{-2}}}{\unit{km}{s^{-1}}}}),
    \end{equation}
    We are given $\dd{N_{\rm HI}} / \dd{\nu}$ in the question and so plugging this in with the optical depth limit we find that the upper limit on the temperature is
    \begin{equation}
        \boxed{ T_{\rm spin} > 824 \unit{K} }
    \end{equation}
}

\question{1d. More absorption}
\answer{
    Repeating parts (b) and (c) but instead using a change of 50\% instead, we find that
    \begin{equation}
        \boxed{ T_{\rm spin} = 12 \unit{K} }
    \end{equation}
}

\question{1e. Spin temperature and absorption}
\answer{
    We note that the trend is \textbf{a higher spin temperature results in a lower detectability of HI-21cm in absorption}. A quick answer to this is that the emissivity is temperature independent whilst the absorptivity is inversely proportional to temperature and thus the rate of absorption declines with increased temperatures.

    More conceptually, an increased spin temperature is the same as an increase excitation temperature, implying that there are a larger number of collisions. This increases number of collisions affects the level populations of HI and increases the rate of emission. As this emission increases, it cancels more absorption and thus decreases the detectability of the absorption.
}

\question{\textbf{2. Broadening}}

\question{2a. Optically Thin Solution}
\answer{
    We have a system of two equations with two unknowns, let's do some rearranging. I'm also going to refer to the FWHM as $w$ for brevity.
    \begin{align}
        \frac{w_1^2}{8 \ln 2} &= \sigma_{v, {\rm turb}}^2 + \frac{k_B T}{M_1},\qquad \frac{w_2^2}{8 \ln 2} = \sigma_{v, {\rm turb}}^2 + \frac{k_B T}{M_2} \\
        \frac{w_1^2}{8 \ln 2} - \frac{k_B T}{M_1} &= \frac{w_2^2}{8 \ln 2} - \frac{k_B T}{M_2} \\
        \frac{w_1^2}{8 \ln 2} - \frac{w_2^2}{8 \ln 2} &= T \qty(\frac{k_B}{M_1} - \frac{k_B}{M_2}) \\
        T &= \frac{w_1^2 - w_2^2}{8 \ln 2} \qty(\frac{k_B}{M_1} - \frac{k_B}{M_2})^{-1}
    \end{align}
    We can plug in $w_1 = 10 \unit{km}{s^{-1}}$, $w_2 = 5 \unit{km}{s^{-1}}$, $M_1 = 1 \unit{amu}$ and $M_2 = 23 \unit{amu}$ to find that the value for an optically thin solution are
    \begin{equation}
        \boxed{ T = 1713 \unit{K},\qquad \sigma_{v, {\rm turb}} = 1.97 \unit{km}{s^{-1}} }
    \end{equation}
}

\question{2b. What if it's not optically thin?}
\answer{
    The FHWM is a larger when it is saturated and so this represents \textbf{an upper limit on the FWHM}.
    \begin{equation}
        \sigma_{v, {\rm turb}}^2 + \frac{k_B T}{M} \le \sigma_v^2
    \end{equation}
    For a similar reason the inferred \textbf{$\sigma_{v, {\rm turb}}$ is an upper limit}. However, if you hold the turbulence fixed and assume the FWHM is an upper limit, this gives that \textbf{the temperature is a lower limit}.
}

>>>>

\newcommand{\psetnum}{9}
\newcommand{\class}{ASTR 541 - Interstellar Medium}

@@

\question{\textbf{1. Molecules}}

\question{Part 1: Rate Equations}
\answer{
    Each rate equation uses the rate coefficient and the number density of each reactant on the left hand side of the equations. The exception is the third where the number density of photons has been accounted for.
    \begin{align}
        \mathcal{R}_{\rm ra} &= k_{\rm ra} n_H n_e \\
        \mathcal{R}_{\rm ad} &= k_{\rm ad} n_H n_{H^{-}} \\
        \mathcal{R}_{\rm pd} &= \zeta_{\rm pd} n_{H^{-}}
    \end{align}
}

\question{Part 2: Rate balance}
\answer{
    Now we can balance these equations and solve for the ratio of ionised to neutral hydrogen.
    \begin{align}
        \mathcal{R}_{\rm ra} &= \mathcal{R}_{\rm ad} + \mathcal{R}_{\rm pd} \\
        k_{\rm ra} n_H n_e &= k_{\rm ad} n_H n_{H^{-}} + \zeta_{\rm pd} n_{H^{-}} \\
        \frac{n_{H^{-}}}{n_H} &= \frac{k_{\rm ra} n_e}{k_{\rm ad} n_H + \zeta_{\rm pd}} \\
        \Aboxed{ \frac{n_{H^{-}}}{n_H} &= 1.36 \times 10^{-11} }
    \end{align}
}

\question{Part 3: Fraction of associated detachment}
\answer{
    We can find the fraction out of total detachment as follows (luckily the H- cancels so we can do this easily!)
    \begin{equation}
        \boxed{ \frac{\mathcal{R}_{\rm ad}}{\mathcal{R}_{\rm ad} + \mathcal{R}_{\rm pd}} = \frac{k_{\rm ad} n_H}{k_{\rm ad} n_H + \zeta_{\rm pd}} = 0.14 }
    \end{equation}
}

\pagebreak

\question{Part 4: Compare to dust}
\answer{
    We can plug in the numbers given in the question to find
    \begin{equation}
        \mathcal{R}_{\rm H^{-}} \equiv k_{\rm ad} \frac{n_{H^{-}}}{n_H} = 1.77 \times 10^{-20} \unit{cm^3}{s^{-1}}
    \end{equation}
    We can take a ratio and find
    \begin{equation}
        \boxed{ \frac{\mathcal{R}_{\rm H^{-}}}{\mathcal{R}_{\rm dc}} \approx 1700 }
    \end{equation}
    So the dust is much stronger!
}

\question{Part 5: What temperature do we need}
\answer{
    Now we can add a temperature dependence and solve for the temperature at which they are equal
    \begin{align}
        1.77 \times 10^{-20} T_2^{0.67} &= 3 \times 10^{-17} \\
        \Aboxed{ T &= 6.6 \times 10^{6} \unit{K} }
    \end{align}
    This is not physically plausible because we'll never get neutral hydrogen in high enough quantities at these temperatures. Our assumption of a constant $n_{\rm H} = 30 \unit{cm^{-3}}$ cannot be valid.
}
